### 2011-04-03 msm2aa.tex
Started msm2aa, made title.tex, modified format.tex

\documentclass[12pt]{article}
\input{./include/format.tex}

% SET THIS STUFF - info for front page.
\newcommand{\nameOfTheModule}{MSM2template}
\newcommand{\nameOfTheAuthor}{Will Ridgers}
\newcommand{\nameOfTheLecturer}{Dr Qianxi Wang}

\begin{document}
\input{./include/title.tex}
\tableofcontents
\pagebreak

\section{Functions of Several Real Variables}
	\begin{defn} 
		For a function $z = f(x,y)$ a vertical section if the graph of $z = f(x,c)$ or $z = f(c,y)$ 
		for some constant $c$. A level curve is the curve $c = f(x,y)$.
	\end{defn}

\section{Partial derivatives and chain rule}

	\begin{defn}
		The function $f: \mathbb{R}^3 \to \mathbb{R}$ is said to have a partial derivative with respect to $x$
		at the point $(x_0, y_0, z_0)$ if there exists a limit
			\[
				\frac{\partial f (x_0, y_0, z_0)}{\partial x} \equiv \lim_{\Delta x \to 0} 
				\frac{f(x_0 + \Delta x, y_0, z_0) - f(x_0, y_0, z_0)}{\Delta x}
			\]
		which is called the partial derivative of $f$ with respect to $x$.
	\end{defn}

	Derivatives with respect to $y$ and $z$ are defined in a similar way.

	\begin{defn}
		The function $f(x,y,z)$ is called differentiable at $(x_0, y_0, z_0)$ if 
		$\Delta f = f(x,y,z) - f(x_0, y_0, z_0)$ can be expressed as
		\[
			\Delta f = f_x(x_0, y_0, z_0) \Delta x + f_y(x_0, y_0, z_0) \Delta y + f_z(x_0, y_0, z_0) \Delta z + o(\rho)
		\]
		where $\Delta x = x - x_0$, $\Delta y = y - y_0$, $\Delta z = z - z_0$ and 
		\[
			\rho = \sqrt{ (\Delta x)^2 + (\Delta y)^2 + (\Delta z)^2 }
		\]
	\end{defn}
	The notation $o(\rho)$ means going to $0$ faster than $\rho$. Note, this is infinitely small so can be ignored.

\section{Vector and vector geometry}
\section{Gradient vector and directional derivative}

\end{document}

### 2011-04-03 msm2aa.tex
Getting somewhere...

@@
 \documentclass[12pt]{article}
-\input{./include/format.tex}
+\input{./include/include.tex}
 
 % SET THIS STUFF - info for front page.
-\newcommand{\nameOfTheModule}{MSM2template}
+\newcommand{\nameOfTheModule}{MSM2Aa}
 \newcommand{\nameOfTheAuthor}{Will Ridgers}
@@
 
+	\subsection{Partial derivatives}
 	\begin{defn}
@@
 
+	\subsection{Differentiable functions}
 	\begin{defn}
@@
 
+	\subsection{Sufficiently smooth}
+	\begin{defn}
+		(Informal) A function is said to be sufficiently smooth if the function and as many of it's partial derivatives as required are continuous where they need to be.
+	\end{defn}
+
+	\subsection{Chain rule}
+	\begin{defn}
+		Chain Rule for two variables. Let $f$ be a function of the variables $x_1, x_2, \cdots , x_n$, $f = f(x_1, x_2, \cdots , x_n)$ 
+		where each $x_j$ is a function of the variables $t_1, t_2, \cdots , t_n$, $x_j = x_j(t_1, t_2, \cdots , t_n)$. Provided
+		that $f$ and $x_j$ are sufficiently smooth, then
+		\[
+			\frac{\partial f}{\partial t_i} = \frac{\partial f}{\partial x_1}\frac{\partial x_1}{\partial t_i} + 
+											\frac{\partial f}{\partial x_2}\frac{\partial x_2}{\partial t_i} + 
+											\cdots +
+											\frac{\partial f}{\partial x_n}\frac{\partial x_n}{\partial t_i}
+		\]
+	\end{defn}
+
 \section{Vector and vector geometry}
+
+	\subsection{Vectors}
+	\begin{defn}
+		A vector has both a magnitude and a direction. Vectors may be denoted as $\mathbf{u}$, $\vec{u}$, or $\underline{u}$.
+		The magnitude of a vector is denoted by 
+		\[
+			|\mathbf{u}| = \sqrt{u_1^2 + u_2^2 + u_3^2}
+		\]
+		A unit vector has length $1$ and in the Cartesian coordinate system $\mathbf{i}$, $\mathbf{j}$, and $\mathbf{k}$ denote the the unit 
+		coordinate vectors along x, y, z-axes respectively.
+	\end{defn}
+
+	\subsection{Scalar (dot) product}
+	\begin{defn}
+		Given two vectors $\mathbf{a} = a_1 \mathbf{i} + a_2 \mathbf{j} + a_3 \mathbf{k}$,
+		$\mathbf{b} = b_1 \mathbf{i} + b_2 \mathbf{j} + b_3 \mathbf{k}$ the scalar (or dot) product of the two is
+		\[
+			\mathbf{a} \cdot \mathbf{b} = |\mathbf{a}||\mathbf{b}|\cos{\theta} = a_1 b_1 + a_2 b_2 + a_3 b_3
+		\]
+		where $\theta$ is the angle between $\mathbf{a}$ and $\mathbf{b}$.
+	\end{defn}
+
+	\subsection{Vector (cross) product}
+	\begin{defn}
+		Given two vectors $\mathbf{a} = a_1 \mathbf{i} + a_2 \mathbf{j} + a_3 \mathbf{k}$,
+		$\mathbf{b} = b_1 \mathbf{i} + b_2 \mathbf{j} + b_3 \mathbf{k}$ the vector (or cross) product of the two is
+		\[
+			\mathbf{a} \times \mathbf{b} = |\mathbf{a}||\mathbf{b}|\sin{\theta}\mathbf{n} =
+			\begin{vmatrix}
+				\mathbf{i} & \mathbf{j} & \mathbf{k} \\
+				a_1 & a_2 & a_3 \\
+				b_1 & b_2 & b_3 
+			\end{vmatrix} 
+		\]
+	\end{defn}
+
+	\subsection{Straight line equation}
+	\begin{defn}
+		The vector equation of a straight line is
+		\[
+			\mathbf{r} = \mathbf{a} + t \mathbf{d}
+		\]
+		for $t \in \mathbb{R}$, where $\mathbf{a}$ is a point on the line, $\mathbf{d}$ is parallel to the line, 
+		and $\mathbf{r}$ is the position vector of a general point on the line.
+	\end{defn}
+
+	\subsection{3 dimensional plane equation}
+	\begin{defn}
+		The vector equation of a plane in 3 dimensions is
+		\[
+			\mathbf{r} \cdot \mathbf{n} = \mathbf{a} \cdot \mathbf{n}
+		\]
+		where $\mathbf{a}$ is the position vector of a point in the plane, $\mathbf{n}$ is a vector normal to
+		the plane and $\mathbf{r}$ is the position vector of a general point in the plane.
+	\end{defn}
+
+	\subsection{Parametric form}
+	A curve $C$ can be expressed using parametric form
+	\[
+		\mathbf{r}(t) = x(t)\mathbf{i} + y(t)\mathbf{j} + z(t)\mathbf{k}
+	\]
+	where $a \leq t \leq b$. Thus the initial point $A$ is $(x(a),y(a),z(a))$ and the terminal point $B$ at $(x(b),y(b),z(b))$.
+	$\mathbf{r}(t)$ is a mapping on the interval $[a,b]$ on the t-axis to the curbe in xyz-space.
+
+	\subsection{Tangent vector of a curve}
+
 \section{Gradient vector and directional derivative}

### 2011-04-03 msm2aa.tex
Nearly done first half of 2Aa

@@
 	\subsection{Tangent vector of a curve}
+	For a curve represented in parametric form (see above), for each $t$ there is a point $\mathbf{r}(t)$ on the curve. The tangent
+	vector of the curve can be found by
+	\begin{align*}
+		\frac{d\mathbf{r}}{dt}& = \lim_{\Delta t \to 0} \frac{\mathbf{r}(t+\Delta t) - \mathbf{r}(t)}{\Delta t} \\
+						& = \lim_{\Delta t \to 0} \frac{x(t+\Delta t) - x(t)}{\Delta t} \mathbf{i} + 
+							  \lim_{\Delta t \to 0} \frac{y(t+\Delta t) - y(t)}{\Delta t} \mathbf{j} + 
+							  \lim_{\Delta t \to 0} \frac{z(t+\Delta t) - z(t)}{\Delta t} \mathbf{k}
+						& = \frac{dx}{dt} \mathbf{i} + \frac{dy}{dt} \mathbf{j} + \frac{dz}{dt} \mathbf{k}
+	\end{align*} 
+	The unit tangent vector $\tau$ is $\tau \equiv \frac{\mathbf{r}_t}{|\mathbf{r}_t|}$. In general $\mathbf{r}' \not{=} 0$ for
+	a sufficiently smooth function.
 
@@
 
+	\subsection{Gradient vector (grad/nabla/del)}
+	\begin{defn}
+		Let $f:\mathbb{R}^3 \to \mathbb{R}$ be a function. If the partial derfivatives of $f$ with respect to $x$, $y$, and $z$
+		at the point $(a,b,c)$ then the gradient vector of $f$ at $(a,b,c)$ is defined to be the vector
+		\[
+			\nabla f(a,b,c) = \frac{\partial f(a,b,c)}{\partial x}\mathbf{i} + 
+								\frac{\partial f(a,b,c)}{\partial y}\mathbf{j} + 
+								\frac{\partial f(a,b,c)}{\partial z}\mathbf{k}
+		\]
+	\end{defn}
+
+	The value of $\nabla f$ at a point is independent of the coordinate system.
+
 \end{document}

### 2011-04-04 msm2aa.tex
50% of msms2aa done. no gfx.

@@
 \section{Functions of Several Real Variables}
+	\subsection{Functions of Several Real Variables}
 		\begin{defn} 
@@
 
-\section{Partial derivatives and chain rule}
+	\subsection{Partial derivatives and chain rule}
 
-	\subsection{Partial derivatives}
+		\subsubsection{Partial derivatives}
 		\begin{defn}
@@
 
-	\subsection{Differentiable functions}
+		\subsubsection{Differentiable functions}
 		\begin{defn}
@@
 
-	\subsection{Sufficiently smooth}
+		\subsubsection{Sufficiently smooth}
 		\begin{defn}
-		(Informal) A function is said to be sufficiently smooth if the function and as many of it's partial derivatives as required are continuous where they need to be.
+			(Informal) A function is said to be sufficiently smooth if the function and as many of it's partial derivatives
+			as required are continuous where they need to be.
 		\end{defn}
 
-	\subsection{Chain rule}
+		\subsubsection{Chain rule}
 		\begin{defn}
@@
 
-\section{Vector and vector geometry}
+	\subsection{Vector and vector geometry}
 
-	\subsection{Vectors}
+		\subsubsection{Vectors}
 		\begin{defn}
@@
 
-	\subsection{Scalar (dot) product}
+		\subsubsection{Scalar (dot) product}
 		\begin{defn}
@@
 
-	\subsection{Vector (cross) product}
+		\subsubsection{Vector (cross) product}
 		\begin{defn}
@@
 
-	\subsection{Straight line equation}
+		\subsubsection{Straight line equation}
 		\begin{defn}
@@
 
-	\subsection{3 dimensional plane equation}
+		\subsubsection{3 dimensional plane equation}
 		\begin{defn}
@@
 
-	\subsection{Parametric form}
+		\subsubsection{Parametric form}
 		A curve $C$ can be expressed using parametric form
@@
 
-	\subsection{Tangent vector of a curve}
+		\subsubsection{Tangent vector of a curve}
 		For a curve represented in parametric form (see above), for each $t$ there is a point $\mathbf{r}(t)$ on the curve. The tangent
@@
 								  \lim_{\Delta t \to 0} \frac{y(t+\Delta t) - y(t)}{\Delta t} \mathbf{j} + 
-							  \lim_{\Delta t \to 0} \frac{z(t+\Delta t) - z(t)}{\Delta t} \mathbf{k}
+								  \lim_{\Delta t \to 0} \frac{z(t+\Delta t) - z(t)}{\Delta t} \mathbf{k} \\
 							& = \frac{dx}{dt} \mathbf{i} + \frac{dy}{dt} \mathbf{j} + \frac{dz}{dt} \mathbf{k}
@@
 
-\section{Gradient vector and directional derivative}
+	\subsection{Gradient vector and directional derivative}
 
-	\subsection{Gradient vector (grad/nabla/del)}
+		\subsubsection{Gradient vector (grad/nabla/del)}
 		\begin{defn}
@@
 
-	The value of $\nabla f$ at a point is independent of the coordinate system.
+		\subsubsection{Key points}
+		\begin{itemize}
+			\item The value of $\nabla f$ at a point is independent of the coordinate system.
+			\item $\nabla f$ is a vector function in n dimensions where f has n variables.
+			\item $\nabla$ is a linear operator, i.e $\nabla (\lambda f + \mu g) = \lambda \nabla f + \mu \nabla g$.
+		\end{itemize}
+
+		\subsubsection{Grad as normal}
+		\begin{thm}
+			Let $f:\mathbb{R}^2 \to \mathbb{R}$ be a sufficiently function, and $(a,b)$ be a point on the level curve $f(x,y)=k$ for some
+			constant $k$. If $\nabla f (a,b) \not{=} 0$, then the vector $\nabla f (a,b)$ is normal to the level curbe $f(x,y)=k$ at the point $(a,b)$.
+		\end{thm}
+
+		\subsubsection{Finding tangent plane to a surface}
+		To find the tangent plane to a surface $z=F(x,y)$ or $G(x,y,z)=k$ at the point $\mathbf{a} = (a,b,c)$:
+		\begin{enumerate}
+			\item Write the surface in the form of the level surface
+				\[
+					f(x,y,z) = F(x,y) - z = 0
+				\]
+				or
+				\[
+					f(x,y,z) = G(x,y,z) - k = 0
+				\]
+			\item Calculate the normal vector to the surface at the point $\mathbf{a}$
+				\[
+					\mathbf{n} = \nabla f ( \mathbf{a} )
+				\]
+			\item The tangent plane is
+				\[
+					\mathbf{r} \cdot \mathbf{n} = \mathbf{a} \cdot \mathbf{n}
+				\]
+		\end{enumerate}
+
+		\subsubsection{Directional derivatives}
+		\begin{defn}
+			Suppose that $f:\mathbb{R}^3 \to \mathbb{R}$ and $\mathbf{u} = u_1 \mathbf{i} + u_2 \mathbf{j} + u_3 \mathbf{k}$ is a unit vector in $\mathbb{R}^3$.
+			The directional derivative of $f$ at $(x,y,z)$ in the direction of $\mathbf{u}$ is defined to be
+			\[
+				D_u f(x,y,z) = \lim_{h \to 0^{+}} \frac{f(x+hu_1, y+hu_2, z+hu_3) - f(x,y,z)}{h}
+			\]
+		\end{defn}
+
+		\begin{lemma}
+			Let $f:\mathbb{R}^3 \to \mathbb{R}$ be a sufficiently smooth, and $\mathbf{u} = u_1 \mathbf{i} + u_2 \mathbf{j} + u_3 \mathbf{k}$ 
+			be a unit vector in $\mathbb{R}^3$. If it exists, then
+			\[
+				\frac{d}{dt} f(a + tu_1, b + tu_2, c + tu_3) | _ t=0 = D_u f(a,b,c)
+			\]
+		\end{lemma}
+
+		\subsubsection{Relation between gradient and directional derivative}
+		\begin{thm}
+			Let $f:\mathbb{R}^3 \to \mathbb{R}$ be a sufficiently smooth, and $\mathbf{u} = u_1 \mathbf{i} + u_2 \mathbf{j} + u_3 \mathbf{k}$ 
+			be a unit vector in $\mathbb{R}^3$. Then
+			\[
+				D_u f(a,b,c) = \mathbf{u} \cdot \nabla f(a,b,c)
+			\]
+		\end{thm}
+
+		\begin{itemize}
+			\item The directional derivative is a scalar function not a vector.
+			\item $D_u$ is linear; $D_u(\lambda f + \mu g) = \lambda D_u f + \mu D_u g$.
+		\end{itemize}
+
+		\subsubsection{Calculating directional derivatives}
+		To calculate the directional derivatibe of a function $f$ in the direction of the vector $\mathbf{u}$:
+		\begin{enumerate}
+			\item Find the unit vector $\hat{\mathbf{u}} = \frac{\mathbf{u}}{|\mathbf{u}|}$
+			\item Find $\nabla f$
+			\item Calculate $D_{\hat{\mathbf{u}}} f = \nabla f \cdot \hat{\mathbf{u}}$
+		\end{enumerate}
+
+		\subsubsection{Steepst ascent theorem}
+		\begin{thm}
+			Let $f$ be a function of several real variables. Suppose that $\nabla f$ exists and is non-zero at the point $\mathbf{a}$. 
+			Then the direction of the vector $\nabla f(\mathbf{a})$ is always the direction of maximum increase of the function $f$ at the point $\mathbf{a}$.
+		\end{thm}
+
+		\subsubsection{Finding paths of maximum increase}
+		To calculate the path of maximum increase of a function $f$ of two variables starting at point $(a,b)$:
+		\begin{enumerate}
+			\item Let the path be $\mathbf{r}(t) = x(t)\mathbf{i} + y(t)\mathbf{j}$
+			\item $\nabla f$ at a point $(x,y)$ is always in the direction of the maximum increase (by Steepst ascent theorem) of the function $f$ at the point,
+					so the tangent vector to the path  $\mathbf{r}'(t) = x'(t)\mathbf{i} + y'(t)\mathbf{j}$ is parallel to $\nabla f = f_x \mathbf{i} + f_y \mathbf{j}$
+			\item Since the two vectors are parallel we have
+					\[
+						x'(t) \mathbf{i} + y'(t) \mathbf{j} = \lambda ( f_x \mathbf{i} + f_y \mathbf{j} ),  \quad \lambda \not{=} 0 \implies \frac{y'(t)}{x'(t)} = \frac{f_y}{f_x}
+					\]
+			\item Hence
+					\[
+						\frac{dy}{dx} = \frac{y'(t)}{x'(t)} = \frac{f_y}{f_x}
+					\]
+			\item Solve for $y$ in terms of $x$ using the boundary conditions $y=b$ when $x=a$.
+		\end{enumerate}
+
+\section{Integration of Functions of Several Real Variables}
+
+	\subsection{Multiple integrals}
 

### 2011-04-04 msm2aa.tex
Updates.
Updates, etc.

@@
 	\subsection{Multiple integrals}
+		\subsubsection{Single integrals}
+
+		\subsubsection{Geometric interpretation of single integrals}
+
+		\subsubsection{Double integrals}
+		\begin{defn}
+			The double integral of a continuous function $f(x,y)$ defined on a closed domain $D$ is defined by
+			\[
+				\iint\limits_D f(x,y) dxdy = \lim_{max|\Delta x_i| \to 0} \lim_{max|\Delta y_j| \to 0} \sum_{j=0}^{m-1} \sum_{i=0}^{n-1} f( \xi_i, \eta_j) \Delta x_i \Delta y_j
+			\]
+			where
+			\[
+				\Delta x_i = x_{i+1}- x_i, \quad \Delta y_j = y_{j+1}- y_j, \quad \xi_i \in [x_i, x_{i+1}], \quad \eta_j \in [x_j, x_{j+1}]
+			\]
+			and the sum for all $(x_i, y_j) \in D$.
+		\end{defn}
+
+		\subsubsection{Geometric interpretation of double integrals}
+
+		\subsubsection{Properties of double integrals}
+		\begin{enumerate}
+			\item If $D$ is a region of $\mathbb{R}^2$, then $\iint\limits_D f(x,y) dxdy$ is the volume of a solid region $S$ lying above domain $D$ in the $xy$-plane and below the surface $z=f(x,y)$.
+			\item $\iint\limits_D 1 dxdy = $ area of $R$.
+			\item If $S$ is a region of $\mathbb{R}^3$ then $\iiint\limits_S 1 dxdydz = $ volume of $S$.
+			\item Suppose $f$ and $g$ are continuous functions, $c$ and $d$ are constants, then
+					\[
+						\iint\limits_D [ cf(x,y) + dg(x,y)] dxdy = c \iint\limits_D f(x,y) dxdy + d \iint\limits_D g(x,y) dxdy
+					\]
+			\item Let $D = D_1 \cup D_2$, where $D_1$ and $D_2$ are non-overlapping regions. If $\iint\limits_D f(x,y)$ dxdy exists, then
+					\[
+						\iint\limits_D f(x,y) dxdy = \iint\limits_{D_1} f(x,y) dxdy + \iint\limits_{D_2} f(x,y) dxdy
+					\]
+		\end{enumerate}
+
+		\subsubsection{Fubini's theorem}
+		\begin{thm}
+			Let $f:\mathbb{R}^2 \to \mathbb{R}$ be continuous on $D$, which is given as
+			\[
+				D = {(x,y): a \leq x \leq b, c(x) \leq y \leq d(x)},
+			\]
+			where $c(x)$ and $d(x)$ are smooth. If $\iint\limits_D f(x,y) dxdy$ exists, then
+			\[
+				\iint\limits_D f(x,y) dxdy = \int\limits_a^b \left( \int\limits_{c(x)}^{d(x)} f (x,y) dy \right) dx = \int\limits_a^b dx \int\limits_{c(x)}^{d(x)} f (x,y) dy
+			\]
+		\end{thm}
+
+		\subsubsection{Geometric interpretation of fubini's theorem}
 

### 2011-04-04 msm2aa.tex
Nearly done 2Aa first draft.

@@
 			\[
-				D = {(x,y): a \leq x \leq b, c(x) \leq y \leq d(x)},
+				D = \{(x,y): a \leq x \leq b, c(x) \leq y \leq d(x)\},
 			\]
@@
 
+		\begin{thm}
+			Let $f:\mathbb{R}^2 \to \mathbb{R}$ be continuous on $D$, which is given as
+			\[
+				D = \{(x,y): c \leq y \leq d, a(y) \leq y \leq b(y)\},
+			\]
+			where $a(y)$ and $b(y)$ are smooth. If $\iint\limits_D f(x,y) dxdy$ exists, then
+			\[
+				\iint\limits_D f(x,y) dxdy = \int\limits_c^d dy \int\limits_{a(y)}^{b(y)} f (x,y) dx
+			\]
+		\end{thm}
+
+		\subsubsection{Fubini's theorem for triple integrals}
+		\[
+			\iiint\limits_S f(x,y,z) dxdydz \equiv \int\limits_{x=a}^{x=b} dx \int\limits_{y=l(x)}^{y=r(x)} dy \int\limits_{z=g(x,y)}^{z=h(x,y)} f(x,y,z) dz
+											\equiv \int\limits_{a}^{b} dx \int\limits_{l(x)}^{r(x)} dy \int\limits_{g(x,y)}^{h(x,y)} f(x,y,z) dz
+		\]
+		The order of integration does not matter. Choose the easiest order.
+
 		\subsubsection{Geometric interpretation of fubini's theorem}
 
+	\subsection{Double integrals in polar coordinates}
+		\subsubsection{Polar coordinates}
+		The relation between polar coordinates and Cartesian coordinates are:
+		\[
+			x = r \cos{\theta}, y = r \sin{\theta}
+		\]
+		or
+		\[
+			r = sqrt{x^2 + y^2}, \tan{\theta} = \frac{y}{x}
+		\]
+		where $r \geq 0$, $0 \leq \theta \leq 2 \pi$ or $-\pi \le \theta \leq \pi$.
+
+		\subsubsection{Double integrals in polar coordinates}
+		For double integrals in polar coordinates
+		\[
+			dA = dxdy = rdrd \theta
+		\]
+		\[
+			\iint\limits_D f(x,y) dxdy = \iint\limits_{D'} f(r,\theta) rdrd \theta
+		\]
+		where $D'$ is the region of the $(r,\theta)$-plane corresponding to the region of $D$ of the $(x,y)$-plane.
+
+		\subsubsection{General case}
+		\begin{defn}
+			The Jacobian of the functions $x(u,v)$ and $y(u,v)$ with respect to the variables $u$ and $v$ is the determinant. Hence,
+			\[
+				\frac{\partial (u,v)}{\partial (x,y)} = 
+				\begin{vmatrix}
+					x_u & x_v  \\
+					y_u & y_v 
+				\end{vmatrix} 
+			\]
+		\end{defn}
+		For double integrals we have
+		\[
+			dA = dxdy = \left| \frac{\partial (u,v)}{\partial (x,y)} \right| dudv
+		\]
+		\[
+			\iint\limits_D f(x,y) dxdy = \iint\limits_{D'} f(x,y) | \frac{\partial (x,y)}{\partial (u,v)} | dudv
+		\]
+		where $D'$ is the region of the $(u,v)$-plane corresponding to the region of $D$ of the $(x,y)$-plane.
+
+	\subsection{Change of variables in triple integrals}
+		\subsubsection{Cylindrical coordinates}
+		Consider the point $P$ with Cartesian coordinates $(x,y,z)$, $Q$ is the projection of $P$ on the $Oxy$ plane. Denote its cylindrical coordinates as $(\rho, \theta, z)$, then:
+		\[
+			\left\{ 
+				\begin{matrix}
+					x = r \cos{\theta} \\
+					y = r \sin{\theta} \\ 
+					z = z
+				\end{matrix} 
+			\right.
+				\iff
+			\left\{ 
+				\begin{matrix}
+					r = \sqrt{x^2 + y^2}  \\
+					\tan{\theta} = \frac{y}{x} \\
+					z = z
+				\end{matrix} 
+			\right.
+		\]
+		where $r \ge 0$, $0 \le \theta < 2 \pi$, $-\infty < z < \infty$.
+
+		\subsubsection{Jacobian for triple integrals}
+		\begin{defn}
+			The Jacobian of the functions $x(u,v,w)$, $y(u,v,w)$, and $z(u,v,w)$ with respect to the variables $u$, $v$, and $w$ is the determinant. Hence,
+			\[
+				\frac{\partial (u,v,w)}{\partial (x,y,z)} = 
+				\begin{vmatrix}
+					x_u & x_v & x_w \\
+					y_u & y_v & y_w \\
+					z_u & z_v & z_w					
+				\end{vmatrix} 
+			\]
+		\end{defn}
+
+		For triple integrals we have that 
+		\[
+			dV = dxdydz = \left| \frac{\partial (x,y,z)}{\partial (u,v,w)} \right| dudvdw
+		\]
+		Hence,
+		\[
+			\iiint\limits_S f(x,y,z) dxdydz = \iiint\limits_{S'} f(x(u,v,w), y(u,v,w), z(u,v,w)) \left| \frac{\partial (x,y,z)}{\partial (u,v,w)} \right| dudvdw
+		\]
+		where $S'$ is the region of the $(u,v,w)$-plane corresponding to the region of $S$ of the $(x,y,z)$-plane.
+
+		\subsubsection{Triple integrals in cylindrical coordinates}
+		For the cylindrical coordinates we have that $x=r \cos{\theta}$, $y=r \sin{\theta}$, and $z=z$.
+		Hence,
+		\[
+			\frac{\partial (x,y,z)}{\partial (r,\theta,z)} = 
+				\begin{vmatrix}
+					x_r & x_\theta & x_z \\
+					y_r & y_\theta & y_z \\
+					z_r & z_\theta & z_z					
+				\end{vmatrix}
+				=
+				\begin{vmatrix}
+					x_r & x_\theta & 0 \\
+					y_r & y_\theta & 0 \\
+					0 & 0 & 1					
+				\end{vmatrix} = r
+		\]
+		Finally giving that
+		\[	
+			dV = dxdydz = rdrd\theta dz
+		\]
+		Hence,
+		\[
+			\iiint\limits_S f(x,y,z) dxdydz = \iiint\limits_{S'} f(r, \theta, z) r drd\theta dz
+		\]
+
+		\subsubsection{Spherical coordinates}
+		Spherical coordinates $(\rho, \theta, \phi)$
+		\[
+			x = \rho \sin{\phi} \cos{\theta}, \quad y = \rho \sin{\phi} \sin{\theta}, \quad z = \rho \cos{\phi}
+		\]
+		where $\rho \ge 0$, $0 \le \theta \le 2 \Pi$, $0 \le \phi \le \Pi$.
+
+		Relation between Cartesian and spherical coordinates
+		\begin{align*}
+			\rho &= \sqrt{x^2 + y^2 + z^2} \\
+			\theta &= \arctan{ \left( \frac{y}{z} \right)} \\
+			\phi &= \arccos{ \left(\frac{z}{\rho} \right)}
+		\end{align*}
+
+		Relation between cylindrical and spherical coordinates
+		\begin{align*}
+			r &= \rho \sin{\phi} \\ 
+			\theta &= \theta \\
+			z &= \rho \cos{\phi}
+		\end{align*}
+
+		\subsubsection{Jacobian of spherical coordinates}
+		Using
+		\[
+			x = \rho \sin{\phi} \cos{\theta}, \quad y = \rho \sin{\phi} \sin{\theta}, \quad z = \rho \cos{\phi}
+		\]
+		we have that
+		\[
+			\frac{\partial (x,y,z)}{\partial (\rho,\phi,\theta)} = \rho^2 \sin{\theta}
+		\]
+
+		\subsubsection{Triple integrals in spherical coordinates}
+		From the above result, we have that
+		\[
+			dV = \rho^2 \sin{\theta} d\rho d\phi d\theta
+		\]
+		Hence
+		\[
+			\iiint\limits_S f(x,y,z) dxdydz = \iiint\limits_{S'} f( \rho, \phi, \theta) \rho^2 \sin{\theta} d\rho d\phi d\theta
+		\]
+
+	\subsection{Line integrals}
+		\subsubsection{Fields}
+		\begin{defn}
+			A scalar function depending on a position is called a scalar field. A vector function depending on a position is called a vector field.
+		\end{defn}
+
+		\subsubsection{Tangential vector}
+		If $\mathbf{r} = (x(t), y(t), z(t)), \quad t \in [t_0, t_1]$ is a parameterisatino of the curve $C$ (preserving orientation) then
+		\begin{itemize}
+			\item The tangential vector of the curve is $\mathbf{T}(t) = \frac{d\mathbf{r}}{dt} = \left( \frac{dx}{dt}, \frac{dy}{dt}, \frac{dz}{dt} \right)$
+			\item It's magnitude is $\left| \mathbf{T} \right| = \left| \frac{d\mathbf{r}}{dt} \right| = \sqrt{ \left( \frac{dx}{dt} \right)^2 + \left( \frac{dy}{dt} \right)^2 + \left( \frac{dz}{dt} \right)^2   }$
+			\item The length $ds$ is $ds = \left| d\mathbf{r} \right| = \left| \frac{d\mathbf{r}}{dt} \right|dt = \left| \mathbf{T} \right| dt$
+		\end{itemize}
+
+		\subsubsection{Line integrals}
+
+		\subsubsection{Line integrals of vector fields}
+		\subsubsection{Other expressions of line integrals}
+		\subsubsection{Work done by a force}
+
 \end{document}

### 2011-04-05 msm2aa.tex
Update

@@
 		\subsubsection{Line integrals}
+		\begin{defn}
+			The line integral of a function $f(\mathbf{r})$ over a curve $C$
+			\[
+				\mathbf{r} = \mathbf{r}(t) = x(t)\mathbf{i} + y(t)\mathbf{j} + z(t)\mathbf{k}, \quad a \le t \le b
+			\]
+			is denoted as $\int\limits_C f(\mathbf{r}) ds$, and is defined by
+			\[
+				\int\limits_C f(\mathbf{r}) ds = \int\limits_a^b f(\mathbf{r}(t))|\mathbf{T}|dt
+			\]
+		\end{defn}
+
+		Some important notes
+		\begin{itemize}	
+			\item $f(\mathbf{r}(t))$ is $f(\mathbf{r})$ along the curve $C$
+			\item $|\mathbf{T}| = \left| \frac{d\mathbf{r}}{dt} \right| = \sqrt{ \left( \frac{dx}{dt} \right)^2 + \left( \frac{dy}{dt} \right)^2 + \left( \frac{dz}{dt} \right)^2 }$
+			\item The right hand size is just a common 1-dimensional definite integrate to $t$.
+		\end{itemize}
 
 		\subsubsection{Line integrals of vector fields}
+		\begin{defn}
+			The line integral of a vector $v(\mathbf{r})$ over a curve $C$
+			\[
+				\mathbf{r} = \mathbf{r}(t) = x(t)\mathbf{i} + y(t)\mathbf{j} + z(t)\mathbf{k}, \quad a \le t \le b
+			\]
+			is denoted as $\int\limits_C \mathbf{v} \cdot d\mathbf{r}$, and is defined by
+			\[
+				\int\limits_C \mathbf{v} \cdot d\mathbf{r} = \int\limits_a^b \mathbf{v}(\mathbf{r}(t)) \cdot \mathbf{T}(t) dt
+			\]
+		\end{defn}
+
+		Some important notes
+		\begin{itemize}
+			\item $\mathbf{v}(\mathbf{r}(t))$ is a vector $\mathbf{v}(\mathbf{r})$ along the curve $C$.
+			\item TODO
+			\item The integrand on the right is a scalar function of $t$
+			\item The right hand size is just a common 1-dimensional definite integrate to $t$
+		\end{itemize}
+
 		\subsubsection{Other expressions of line integrals}

### 2011-04-05 msm2ba.tex
Started 2ca (Will) and 2ba (Joao)

\documentclass[12pt]{article}
\input{./include/include.tex}

% SET THIS STUFF - info for front page.
\newcommand{\nameOfTheModule}{MSM2Ba}
\newcommand{\nameOfTheAuthor}{Jo\~{a}o Cho\c{c}a}
\newcommand{\nameOfTheLecturer}{Dr Neal Bez}

\begin{document}
\input{./include/title.tex}
\tableofcontents
\pagebreak

\section{Assumed Knowledge}
	\subsection{Intervals}
		\begin{defn}
			An interval is a subset of $\mathbb{R}$ taking one of the following forms, where $a$ and $b$ are arbitrary real numbers and $a < b$.

			\begin{itemize}
				\item $(a,b) := \{x \in \mathbb{R}: a < x < b\}$ (this is called an open interval)
				\item $(a,b] := \{x \in \mathbb{R}: a < x \leq b\}$	
				\item $[a,b) := \{x \in \mathbb{R}: a \leq x < b\}$		
				\item $[a,b] := \{x \in \mathbb{R}: a \leq x \leq b\}$ (this is called a closed interval)
				\item $(a,\infty) := \{x \in \mathbb{R}: x > a\}$		
				\item $[a,\infty) := \{x \in \mathbb{R}: x \geq a\}$		
				\item $(-\infty,b) := \{x \in \mathbb{R}: x < b\}$		
				\item $(-\infty,b] := \{x \in \mathbb{R}: x \leq b\}$
				\item $(-\infty, \infty) := \mathbb{R}$
				\item The interval $(0, \infty)$ will also be denoted by $\mathbb{R}^+$.
			\end{itemize}

		\end{defn}

\end{document}

### 2011-04-05 msm2ca.tex
Started 2ca (Will) and 2ba (Joao)

\documentclass[12pt]{article}
\input{./include/include.tex}

% SET THIS STUFF - info for front page.
\newcommand{\nameOfTheModule}{MSM2Ca - Linear Algebra}
\newcommand{\nameOfTheAuthor}{Will Ridgers}
\newcommand{\nameOfTheLecturer}{Dr Sergey Shpectorov}

\begin{document}
\input{./include/title.tex}
\tableofcontents
\pagebreak

\section{Introduction}
	\subsection{Fields}
	The set of scalars will often be $\mathbb{R}$, $\mathbb{C}$, or $\mathbb{Q}$. These sets are mathematical structures called fields. Other examples of fields would be
	\begin{itemize}
		\item $\mathbb{Z}_p$ (integers of modulo $p$) where $p$ is a prime integer.
		\item $\mathbb{F}_q$ where $q = p^q \not{=} 1$ is a prime power. This is a unique finite field.
	\end{itemize}

	\subsection{Working with finite fields}
	In $\mathbb{F}_p = \mathbb{Z}_p$ you can simply do the operations in modulo $p$.
	\[
		x \oplus y := x + y mod p,
	\]
	\[
		x \otimes y := xy mod p
	\]
	Subtration and division are subsituted with addition and multiplication:
	\[
		x - y := x + (-y),
	\]
	\[
		\frac{x}{y} := xy^{-1}
	\]

	\subsubsection{Operations for $\mathbb{F}_4$}
	For any finite field you can use precomupted addition/multiplication tables.
	\[
		\begin{array}{c | c c c c}
			+ & 0 & 1 & a & b \\
			\hline
			0 & 0 & 1 & a & b \\
			1 & 1 & 0 & b & a \\ 
			a & a & b & 0 & 1 \\
			b & b & a & 1 & 0
		\end{array}
		\quad
		\begin{array}{c|cccc}
			\cdot & 0 & 1 & a & b \\
			\hline
			0 & 0 & 0 & 0 & 0 \\
			1 & 0 & 1 & a & b \\
			a & 0 & a & b & 1 \\
			b & 0 & b & 1 & a
		\end{array}
	\]
	In $\mathbb{F}_4$, note that $-0 = 0$, $-1 = 1$, $-a = a$, and $-b = b$.

\section{Vector Spaces}
	\begin{defn}
		A vector space over the field of scalars $\mathbb{F}_4$ is a set $V$, whose elements are vectors, together with two binary operations addition and multiplication. The following axioms hold for addition
		\begin{itemize}
			\item Associative: $\forall a,b,c \in V, \quad a + ( b + c ) = (a + b) + c$.
			\item Existance of zero element: $\forall a \in V, \quad \exists 0 \in V$ such that $a + 0 = a$.
			\item Existance of inverse element: $\forall a \in V, \quad \exists (-a) \in V$ such that $a + (-a) = 0$.
			\item Commutative: $\forall a,b \in V, \quad a + b = b + a$.
		\end{itemize}
		And the axioms for multiplication are
		\begin{itemize}
			\item Associative: $\forall a,b \in \mathbb{F}_4, \quad a(bc) = (ab)c$.
			\item Distributive: $\forall a,b,c \in \mathbb{F}_4$ then
			\item Existance of inverse element:
		\end{itemize}
	\end{defn}

\section{Subspaces}
\section{Bases and dimension}
\section{Linear mappings}
\section{Isomorphisms}
\section{Matrix of linear mapping}
\section{Linear transformations}
\section{Eigenvectors}

\end{document}

### 2011-04-05 msm2ca.tex
2ca update.

@@
 		\begin{itemize}
-			\item Associative: $\forall a,b,c \in V, \quad a + ( b + c ) = (a + b) + c$.
-			\item Existance of zero element: $\forall a \in V, \quad \exists 0 \in V$ such that $a + 0 = a$.
-			\item Existance of inverse element: $\forall a \in V, \quad \exists (-a) \in V$ such that $a + (-a) = 0$.
-			\item Commutative: $\forall a,b \in V, \quad a + b = b + a$.
+			\item Associative: $\forall u,v,w \in V, \quad u + ( v + w ) = (u + v) + w$
+			\item Existance of zero element: $\forall u \in V, \quad \exists 0 \in V$ such that $u + 0 = u$
+			\item Existance of inverse element: $\forall u \in V, \quad \exists (-u) \in V$ such that $u + (-u) = 0$
+			\item Commutative: $\forall u,v \in V, \quad u + v = v + u$
 		\end{itemize}
@@
 		\begin{itemize}
-			\item Associative: $\forall a,b \in \mathbb{F}_4, \quad a(bc) = (ab)c$.
-			\item Distributive: $\forall a,b,c \in \mathbb{F}_4$ then
-			\item Existance of inverse element:
+			\item Associative: $\forall a,b \in \mathbb{F}_4, v \in V, \quad a(bv) = (ab)v$.
+			\item Distributive: $\forall a,b \in \mathbb{F}_4, v,u \in V$ then
+					\[
+						a(u+v) = au + av,
+					\]
+					\[
+						(a+b)v = av + bv
+					\]
+			\item Existance of identity element: $1v = v \forall v \in V$
 		\end{itemize}
@@
 
+	\subsection{Basic properties}
+	\begin{thm}
+		Suppose that $V$ is a vector space over a field of scalars $\mathbb{F}$. Then the following axioms hold:
+		\begin{itemize}
+			\item $\forall u,v,w \in V$ if $u + v = u + w$ then $v=w$ (cancellation of sums)
+			\item $0v = 0, \quad \forall v \in V$
+			\item $(-1)v = -v, \quad \forall v \in V$
+			\item $a0 = 0, \quad \forall a \in \mathbb{F}$
+		\end{itemize}
+	\end{thm}
+
+	\begin{thm}
+		Suppose $V$ is a vector space over $\mathbb{F}$, and $a \in \mathbb{F}, \quad u,v \in V$
+		\begin{itemize}
+			\item If $av = 0$ then either $a = 0$ or $v = 0$
+			\item $a \not{=} 0$, if $au = av$ then $u=v$
+			\item $u \not{=} 0$, if $au = bu$ then $a=b$
+		\end{itemize}
+	\end{thm}
+
+	\subsection{Set of functions}
+	For two sets $X$ and $Y$, the set of functions (mappings) from $X$ to $Y$ is denoted by $Y^X$. If $X$ and $Y$ are finite sets with
+	$n$ and $m$ elements, then size of $Y^X$, is exactly $m^n$.
+
 \section{Subspaces}
+
+	\subsection{Vector subspaces}
+	\begin{defn}
+		Let $V$ be a vector space defined over the field of scalars $\mathbb{F}$ and $U \subset V$, $U$ is subspace if it is a vector space over $\mathbb{F}$ for the operations inherited from $V$. 
+		To be a subspace $U$ must
+		\begin{itemize}
+			\item Contain the zero vector 0
+			\item Closed with respect to addition
+			\item Closed with respect to multiplication
+			\item Closed with respect to taking negative vectors
+		\end{itemize}
+	\end{defn}
+
+	\begin{thm}
+		A nonempty subset $U$ of a vector space $V$ is a subspace if and only if $U$ is closed with respect to addition and scalar multiplication.
+	\end{thm}
+
+	\subsection{Intersection of subspaces}
+	\begin{thm}
+		If $\{ U_i | i \in I\}$ is a collection of subspaces of a vector space $V$ then the intersection of $W := \cup_{i \in I} U_i$ is again a subspace of $V$.
+	\end{thm}
+
 \section{Bases and dimension}

### 2011-04-06 msm2ca.tex
Commit because I can.
Finished subspaces.

@@
 		\begin{itemize}
-			\item Contain the zero vector 0
+			\item Contain the zero vector $0$
 			\item Closed with respect to addition
@@
 
+	\subsection{Subspace spanned by a set of vectors}
+	\begin{thm}
+		Suppose $X$ is a set of vectors from a vector space $V$. The subspace spanned by $X$ is the smallest subspace $W$ of $V$ containing $X$. We write that $W = \left< X \right>$.
+	\end{thm}
+
+	\subsection{Linear combinations}
+	\begin{defn}
+		A linear combination of the vectors from $X$ is an expression of the form $w = \sum_{i \in I} a_i u_i$ where $a_i \in \mathbb{F}$ (scalars) and $u_i \in X$ (vectors).
+		If the set $X$ is infinite, them only finitely many $a_i$ are allowed to be nonzero. $w$ will of course be a vector, referred to as a linear combination of the $u_i$'s.
+	\end{defn}
+
+	\subsection{Subspace spanned by vectors}
+	\begin{thm}
+		Suppose $V$ is a vector space over $\mathbb{F}$, and $X = \{ u_i | i \in I \}$ is a set of vectors in $V$. Then $W=\left< X \right>$ consists of all vectors $w$ that are linear
+		combinations of the vectors from $X$.
+	\end{thm}
+
+	\subsection{Is a vector in the subspace?}
+	From above theorem, to check a given vector is contained in the subspace $W=\left< X \right>$, just check $w$ is a linear combination of the vectors $u_i$ from $X$.
+
 \section{Bases and dimension}
+	\begin{defn}
+	\end{defn}
+
 \section{Linear mappings}

### 2011-04-06 msm2ba.tex
Syncing Joao's updates.

@@
 		\begin{defn}
-			An interval is a subset of $\mathbb{R}$ taking one of the following forms, where $a$ and $b$ are arbitrary real numbers and $a < b$.
+			An interval is a subset of $\mathbb{R}$ taking one of the following forms, where ${a}$ and ${b}$ 					are arbitrary real numbers and ${a}$ < ${b}$:
 
@@
 			\end{itemize}
+		\end{defn}
+
+	\subsection{Functions}
+		\begin{defn}
+			Suppose that X $\subseteq \mathbb{R}$.
+A real function $f : X \to \mathbb{R}$ is a a rule which assigns to 				every real number $x \in X$ a unique real number $y \in \mathbb{R}$.
+If this number $y \in 						\mathbb{R}$ corresponds to the number $x \in \mathbb{R}$, then we write $y = f(x)$.
+		\end{defn}
+
+		\begin{defn}
+			The set X where $f$ is defined is called the domain of $f$. We call the set $f(X) = \{f(x) : x \in X\}$ the 						image or range or image of $f$. The graph of $f$ is defined to be
+							\[
+							\{(x, f(x)) \in \mathbb{R}^2 : x \in X\}
+							\]
 
@@
 
+		\subsubsection{Behaviour of Functions at Inifinity}
+			\begin{defn}
+				Suppose that $f$ is a real function.
+Then we say that "$f(x)$ tends to infinity as $x$ tends to 							infinity" and we write $f(x) \to \infty$ as $x \to \infty$ or $lim_{x \to \infty} f(x) = \infty$, 						if, given any A in $\mathbb{R}^+$, $\exists$ K in $\mathbb{R}$ (which may depend on A) 						such that
+					\[
+						f(x) > A \mbox{    whenever    } x > K.
+					\]
+
+			\end{defn}
+
+			\begin{defn}
+				Suppose that $f$ is a real function.
+Then we say that "$f(x)$ tends to minus infinity as $x$ tends to 					infinity" and we write $f(x) \to -\infty$ as $x \to \infty$ or $\lim_{x \to \infty} f(x) = 							-\infty$, if, given any A in $\mathbb{R}^+$, $\exists$ K in $\mathbb{R}$ (which may 							depend on A) such that
+					\[
+						f(x) < -A \mbox{    whenever    } x > K.
+					\]
+
+				It then follows that "f(x) $\to -\infty$ as $x \to \infty$" is logically equivalent to the statement "-f(x) 					$\to -\infty$ as $x \to \infty$"
+			\end{defn}
+
+			{\bf{Remark}} : Note that $|y - \ell| < \epsilon$   if and only if   $\ell - \epsilon < y < \ell + \epsilon$.
+
+			\begin{defn}
+				Let $f$ be a real function and $\ell \in \mathbb{R}$.
+Then we say "f(x) tends to $\ell$ as x tends to 					infinity", and we write $f(x) \to \ell$ as $x \to \infty$, or $\lim_{x \to \infty} = \ell$, if, given any 					$\epsilon$ in $\mathbb{R}^+$, $\exists$ K in $\mathbb{R}$ (which may depend on $\epsilon$) such 					that
+						\[
+							|f(x) - \ell| < \epsilon \mbox{    whenever    } x > K.
+						\]
+
+				In other words, this means that we can ensure |$f(x) - /ell$| remains smaller than any prescribed 					positive real number, provided we take x sufficiently large.
+			\end{defn}
+
+			{\bf{Remark}} : We say such an $\ell$ exists if the limit is defined. If $f(x) \to \infty$ as $x \to \infty$, 							then we say the limit does not exist.
+
+			\begin{thm}
+				Suppose that $f(x) \to \infty$ as $x \to \infty$. Then
+						\[
+							\lim_{x \to \infty}{\frac{1}{f(x)}} = 0
+						\]
+			\end{thm}
+
+			We can now link the ideas in definitions 4, 5 and 6 to produce a definition for what happens when x goes to 				minus infinity.
+
+			\begin{defn}
+				Let $f$ be a real function and $\ell \in \mathbb{R}$. Then
+				\begin{enumerate}
+					\item $f(x) \to \infty$ as $x \to -\infty$, if, given any A in $\mathbb{R}^+$, $\exists$ K in 							$\mathbb{R}$ (which may depend on A) such that
+								\[
+									f(x) > A \mbox{    whenever    } x < K.
+								\]
+					\item $f(x) \to -\infty$ as $x \to -\infty$, if, given any A in $\mathbb{R}^+$, $\exists$ K in 							$\mathbb{R}$ (which may depend on A) such that
+								\[
+									f(x) < -A \mbox{    whenever    } x < K.
+								\]			
+					\item $f(x) \to \ell$ as $x \to -\infty$, if, given any $\epsilon$ in $\mathbb{R}^+$, $\exists$ 						K in $\mathbb{R}$ (which may depend on $\epsilon$) such that
+						\[
+							|f(x) - \ell| < \epsilon \mbox{    whenever    } x < K.
+						\]
+				\end{enumerate}
+			\end{defn}
+
+		\subsubsection{Limits}
+			\begin{thm}
+				(Algebra of Limits) Suppose that
+					\[
+						\lim_{x \to \infty}{f(x)} = \ell     \mbox{     and     }     \lim_{x \to \infty}{g(x)} = m
+					\]
+				where $\ell, m \in \mathbb{R}$. Then the following statements hold:
+					\begin{enumerate}
+						\item $\lim_{x \to \infty}{(f(x) + g(x))} = \ell + m$;
+						\item $\lim_{x \to \infty}{(f(x) - g(x))} = \ell - m$;
+						\item $\lim_{x \to \infty}{(f(x)g(x))} = \ell m$;
+
+				And, if we also suppose that $m \neq 0$, then
+
+						\item $\lim_{x \to \infty}{\frac{f(x)}{g(x)}} = \frac{\ell}{m}$;
+					\end{enumerate}
+
+				Analogous statements hold for limits as $x \to -\infty$.
+			\end{thm}
+
+			\begin{thm}
+				(Sandwich Theorem) Let f, g and h be real functions defined on (R, $\infty$) for some R $\in 						\mathbb{R}$. Suppose that
+						\[
+							f(x) \leq g(x) \leq h(x)
+						\]
+				for all $x \in (R, \infty)$ and
+						\[
+							\lim_{x \to \infty}{f(x)} = \ell \mbox{    and    } \lim_{x \to \infty}{h(x)} = \ell
+						\]
+				Then $\lim_{x \to \infty}{g(x)} = \ell$.
+
+				An analogous statement is true for limite as $x \to -\infty$.
+			\end{thm}
+
+			\begin{defn}
+				(Right-Hand Limit) Suppose that f is defined on ($a$, $a$ + R) for some $R \in \mathbb{R}^+$.
+							Then we say "f(x) tends to $\ell$ as $x$ tends to $a$ from above" if, given any $\epsilon$ in 					$\mathbb{R}^+$, $\exists \delta$ in $\mathbb{R}^+$ such that
+					\[
+						|f(x) - \ell| < \epsilon \mbox {    whenever    } a < x < a + \delta.
+					\]
+
+				In symbols, we write $f(x) \to \ell$ as $x \to a+$, or
+					\[
+						\lim_{x \to a+}{f(x)} = \ell
+					\]
+
+				The symbol $a+$ indicates that the limit is taken from the right, or from above.
+			\end{defn}
+
+			\begin{defn}
+				(Left-Hand Limit) Suppose that f is defined on ($a$ - R, $a$) for some $R \in \mathbb{R}^+$.
+Then 					we say "f(x) tends to $\ell$ as $x$ tends to $a$ from below" if, given any $\epsilon$ in 							$\mathbb{R}^+$, $\exists \delta$ in $\mathbb{R}^+$ such that
+					\[
+						|f(x) - \ell| < \epsilon \mbox {    whenever    } a - \delta < x < a.
+					\]
+
+				In symbols, we write $f(x) \to \ell$ as $x \to a-$, or
+					\[
+						\lim_{x \to a-}{f(x)} = \ell
+					\]
+
+				The symbol $a-$ indicates that the limit is taken from the left, or from below.
+			\end{defn}
+
+			\begin{defn}
+				($\epsilon-\delta$ definition of a limit at a point) Suppose that f is defined on $(a - R, a) \cup (a, a + 					R)$, for some R $\in \mathbb{R}^+$.
+Then we say "f(x) tends to $\ell$ as $x$ tends to 							$a$" if, given any $\epsilon$ in $\mathbb{R}^+, \exists \delta$ in $\mathbb{R}^+$ such 						that
+					\[
+						|f(x) - \ell| < \epsilon \mbox{    whenever    } 0 < |x - a| < \delta
+					\]
+				Using symbols, we write f(x) $\to \ell$ as $x \to a$, or
+					\[
+						\lim_{x \to a}{f(x)} = \ell
+					\]
+			\end{defn}
+
+			{\bf{Remarks}} : 
+			\begin{enumerate}
+				\item The assumptions made for the last 3 definitions do not preclude the possibility that f is 						defined on a larger set than the set considered ($\mathbb{R}$).
+However, the conditions on 						$x$ strictly exclude the possibility that $x = a$.
+In particular, $\lim_{x \to a}{f(x)}$, 							if it exists, is independent of $f(a)$, and it does not matter whether $f(a)$ is defined.
+
+				\item Sometimes we call the set $\{x \in \mathbb{R} : 0 < |x - a| < \delta \}$ a punctured 							neighbourhood of a.
+
+				\item It follows from the definitions that $\lim_{x \to a}{f(x)} = \ell$ if and only if both $\lim_{x \to 						 a-}{f(x)} = \ell$ and $\lim_{x \to a+}{f(x)} = \ell$.
+
+				\item Quite often both  $\lim_{x \to a-}{f(x)}l$ and  $\lim_{x \to a+}{f(x)}$ exist, but are different. 						This is a common reason for  $\lim_{x \to a}{f(x)}$ to fail to exist.
+			\end{enumerate}
+
+			{\bf{Recall that}} if ($a_n$) is a sequence of real numbers that converges to a limit $a \in \mathbb{R}$, 							then we write
+				\[
+					\lim_{n \to \infty}{a_n} = a.
+				\]
+
+			\begin{thm}
+				(A theorem connecting sequences and functions - Theorem 4.6) Suppose that f is a real function, 						defined on (a - R, a) $\cup$ (a, a + R) for some R in $\mathbb{R}^+$.
+Then $\lim_{x \to a}						{f(x)} = \ell$ if and only if $\lim_{n \to \infty}{f(a_n)} = \ell$ for all sequences ($a_n$) such 						that $\lim_{n \to \infty}{a_n} = a$ and $a_n \neq a$ for all $n \in \mathbb{N}$.
+			\end{thm}
+
+			Then, it follows that we can use the Algebra of Limits and the Sandwich Theorem on limits to a point as 						opposed to functions that tend to infinity only.
+
+			\begin{thm}
+				(Algebra of Limits) Suppose that f and g are real functions, whose domains both include a set of the 						form (a - R, a) $\cup$ (a, a + R) for some R in $\mathbb{R}^+$. Suppose also that
+					\[
+						\lim_{x \to a}{f(x)} = \ell     \mbox{     and     }     \lim_{x \to a}{g(x)} = m
+					\]
+				where $\ell, m \in \mathbb{R}$. Then the following statements hold:
+					\begin{enumerate}
+						\item $\lim_{x \to a}{(f(x) + g(x))} = \ell + m$;
+						\item $\lim_{x \to a}{(f(x) - g(x))} = \ell - m$;
+						\item $\lim_{x \to a}{(f(x)g(x))} = \ell m$;
+
+				And, if we also suppose that $m \neq 0$, then
+
+						\item $\lim_{x \to a}{\frac{f(x)}{g(x)}} = \frac{\ell}{m}$;
+					\end{enumerate}
+			\end{thm}
+
+			\begin{thm}
+				(Sandwich Theorem) Suppose that f, g and h are real functions, whose domains all include a set of 						the form (a - R, a) $\cup$ (a, a + R) for some R in $\mathbb{R}^+$. Suppose also that
+						\[
+							f(x) \leq g(x) \leq h(x)
+						\]
+				for all $x \in (a - R, a) \cup (a, a + R))$ and
+						\[
+							\lim_{x \to a}{f(x)} = \ell \mbox{    and    } \lim_{x \to a}{h(x)} = \ell
+						\]
+				Then 
+						\[
+							\lim_{x \to a}{g(x)} = \ell.
+						\]
+			\end{thm}
+
+		\subsubsection{Fundamental Trigonometric Limits}
+
+			{\bf{Remark}}	The $\epsilon-\delta$ definition of a limit can be used to prove that if a, b, c $\in 							\mathbb{R}$ and b $\neq 0$, then
+						\[
+						\lim_{x \to a}{f(bx)} = \lim_{x \to ba}{f(x)}
+						\]			
+				and
+						\[
+						\lim_{x \to a}{f(x + c)} = \lim_{x \to a + c}{f(x)}
+						\]
+
+			{\bf{Important Note}} Unless a question specifically requires the computation of these limits, they are 						standard results and may be used as such.
+In all of them, $a, b, c \in \mathbb{R}, b \neq 0$, and 					$n \in \mathbb{N}$:
+					\begin{enumerate}
+						\item $\lim_{x \to a}{x^n} = a^n$
+						\item $\lim_{x \to a}{f(bx + c)} = \lim_{x \to ab + c}{f(x)}$ (follows from remark 									above)
+						\item $\lim_{x \to 0}{\cos x} = 1$
+						\item $\lim_{x \to 0}{\sin x} = 1$
+						\item $\lim_{x \to 0}{\frac{\sin x}{x}} = 1$
+					\end{enumerate}
+
+		\subsubsection{Continuity}
+
+			\begin{defn}
+				(Continuity at a point) Let $f : X \to \mathbb{R}$ be a real function, and suppose that $a \in X$.
+							Then $f$ is continuous at $a$ if, given any $\epsilon \in \mathbb{R}^+, \exists \delta \in 							\mathbb{R}^+$ such that
+						\[
+							|f(x) - f(a)| < \epsilon \mbox{    whenever    } x \in X \mbox {    and    } 										|x - a| < \delta.
+						\]
+			\end{defn}
+
+			{\bf{Remarks}} :
+					\begin{enumerate}
+						\item If f($x$) is defined for all $x$ in a neighbourhood of a, that is, $(a - R, a + R) 								\subseteq X$ for some $R \in \mathbb{R}^+$, then $f$ is continuous at 									$a$ if and only if
+							\[
+								\lim_{x \to a}{f(x)} = f(a)
+							\] 
+						\item If $f$ is defined on an interval $[a,b]$, where $a,b \in \mathbb{R}$, then $f$ 								is continuous at the endpoint $a$ if and only if
+							\[
+								\lim_{x \to a+}{f(x)} = f(a)
+							\] 
+							similarly, $f$ is continuous at $b$ if and only if
+							\[
+								\lim_{x \to b-}{f(x)} = f(b)
+							\] 
+						\item If $f(a)$ is defined, but $f(x)$ is not defined for any other $x$ in (a - R, a + R) 							for some $R \in \mathbb{R}^+$, then $f$ is continuous at a.
+					\end{enumerate}
+
+			\begin{defn}
+				(Continuity) Let $f : X \to \mathbb{R}$ be a real function. Then $f$ is continuous if $f$ is 								continuous at every point of X.
+			\end{defn}
+
+			{\bf{Remark}}: A function $f : [a,b] \to \mathbb{R}$ is continuous if and only if
+							\[
+								\lim_{x \to c}{f(x)} = f(c) \mbox {   for all   } c \in [a,b]
+							\] 
+						and
+							\[
+								\lim_{x \to a+}{f(x)} = f(a)  \mbox{    and    } \lim_{x \to b-}{f(x)} = 									 f(b)
+							\]
+
+			\begin{thm}
+			(Algebra of Continuous Functions) Suppose that $f : X \to \mathbb{R}$ and $g : X \to \mathbb{R}$ are 						continuous at a point $a \in X$. Then the following statements hold:
+				\begin{enumerate}
+					\item $f + g$ is continuous at $a$;
+					\item $f - g$ is continuous at $a$;
+					\item $fg$ is continuous at $a$;
+
+				And, if we also suppose that $g(a) \neq 0$, then
+
+					\item $\frac{f}{g}$ is continuous at $a$;
+				\end{enumerate}
+			\end{thm}
+
+			\begin{thm}
+				(Corollary to above) Suppose that $f : X \to \mathbb{R}$ and $g : X \to \mathbb{R}$ are 							continuous. Then the following statements hold:
+				\begin{enumerate}
+					\item $f + g$ is continuous;
+					\item $f - g$ is continuous;
+					\item $fg$ is continuous;
+
+				And, if we also suppose that $g(x) \neq 0$ for all $x \in X$, then
+
+					\item $\frac{f}{g}$ is continuous;
+				\end{enumerate}
+			\end{thm}
+
+			\begin{defn}
+				Suppose that $f : X \to Y$ and $g : Y \to Z$ are functions. Then the composed function or 							composition $g \circ f : X \to Z$ is defined by
+					\[
+						g \circ f(x) = g(f(x)) \mbox{   for all    } x \in X.
+					\]
+			\end{defn}
+
+			\begin{thm}
+				(Functions composed of continuous functions are continuous) Suppose that $f : X \to Y$ and $f : Y 						\to Z$ are functions, that $f$ is continuous at a point $a \in X$ and that $g$ is continuous at 					the point $f(a) \in Y$.
+Then the composed function $g \circ f : X \to Z$ is continuous at $a$.
+
+				Similarly, as a corollary, if X and Y are intervals, and if $f : X \to Y$ and $f : Y \to \mathbb{R}$ are 						continuous functions.
+Then the composed function $g \circ f : X \to \mathbb{R}$ is 							continuous.
+			\end{thm}
+
+		\subsubsection{The Intermediate Value Theorem}
+
+			\begin{lemma}
+				Suppose that $f : [a,b] \to \mathbb{R}$ is continuous and $\lambda \in \mathbb{R}$. If $f(w) > 							\lambda$ for some $w \in [a,b]$, then $\exists \delta$ in $\mathbb{R}^+$ such that
+					\[
+						f(x) > \lambda \mbox{   whenever   } x \in [a,b] \mbox{   and   } |x-w| < \delta
+					\]
+				The analogous result holds if the inequality signs > are replaced by <.
+			\end{lemma}
+
+			\begin{lemma}
+				Suppose that $f : [a,b] \to \mathbb{R}$ is continuous and $\lambda \in \mathbb{R}$. If $f(x) \geq 							\lambda$ for all $x \in (a,b)$, then
+					\[
+						f(a) \geq \lambda \mbox{   and   } f(b) \geq \lambda
+					\]
+				The same holds if the inequality signs $\geq$ are replaced by $\leq$.
+			\end{lemma}
+
+			\begin{thm}
+				(The Intermediate Value Theorem) Suppose that $f : [a,b] \to \mathbb{R}$ is continuous, that $f(a) 					\neq f(b)$, and that $\lambda$ lies strictly between $f(a)$ and $f(b)$.
+Then there exists 							$c$ in $(a,b)$ such that
+					\[
+						f(c) = \lambda
+					\]
+			\end{thm}
+
+			{\bf{Remark}} : The continuity hypothesis in the statement of the IVT is essential.
+
+			\begin{thm}
+				(Fixed Point Theorem) If $f : [a,b] \to [a,b]$ is continuous, then $\exists c \in [a,b]$ such that $f(c) 					= c$.
+			\end{thm}
+
+		\subsubsection{Continuous Functions on Closed Bounded Intervals}
+
+			{\bf{Remarks}} : In MSM1B, we stated in proved the following theorems:
+				\begin{enumerate}
+					\item
+						\begin{thm}
+							(Least Upper Bound Axiom) Any bounded set X of real numbers has a least 									upper bound.
+						\end{thm}
+					\item
+						\begin{thm}
+							(Monotone Convergence Theorem) Any monotone bounded sequence 										$\{x_n\}$ of real numbers is convergent.
+						\end{thm}
+					\item
+						\begin{thm}
+							(Bolzano-Weierstrass) Any bounded sequence $\{x_n\}$ of real numbers has 								 a convergent subsequence.
+						\end{thm}
+				\end{enumerate}
+
+			\begin{defn}
+				A function $f : X \to \mathbb{R}$ is bounded if there is a number $R \in \mathbb{R}^+$ such that
+					\[
+						|f(x)| \leq R \mbox{   for all   } x \in X.
+					\]
+			\end{defn}
+
 \end{document}

### 2011-04-06 msm2ca.tex
update.

@@
 	\]
-	Subtration and division are subsituted with addition and multiplication:
+	Subtraction and division are substituted with addition and multiplication:
 	\[
@@
 	\subsubsection{Operations for $\mathbb{F}_4$}
-	For any finite field you can use precomupted addition/multiplication tables.
+	For any finite field you can use precomputed addition/multiplication tables.
 	\[
@@
 	\begin{defn}
-		A vector space over the field of scalars $\mathbb{F}_4$ is a set $V$, whose elements are vectors, together with two binary operations addition and multiplication.
-The following axioms hold for addition
+		A vector space over the field of scalars $\mathbb{F}_4$ is a set $V$, whose elements are 
+		vectors, together with two binary operations addition and multiplication. The following axioms hold for addition
 		\begin{itemize}
 			\item Associative: $\forall u,v,w \in V, \quad u + ( v + w ) = (u + v) + w$
-			\item Existance of zero element: $\forall u \in V, \quad \exists 0 \in V$ such that $u + 0 = u$
-			\item Existance of inverse element: $\forall u \in V, \quad \exists (-u) \in V$ such that $u + (-u) = 0$
+			\item Existence of zero element: $\forall u \in V, \quad \exists 0 \in V$ such that $u + 0 = u$
+			\item Existence of inverse element: $\forall u \in V, \quad \exists (-u) \in V$ such that $u + (-u) = 0$
 			\item Commutative: $\forall u,v \in V, \quad u + v = v + u$
@@
 					\]
-			\item Existance of identity element: $1v = v \forall v \in V$
+			\item Existence of identity element: $1v = v \forall v \in V$
 		\end{itemize}
@@
 	\begin{defn}
-		Let $V$ be a vector space defined over the field of scalars $\mathbb{F}$ and $U \subset V$, $U$ is subspace if it is a vector space over $\mathbb{F}$ for the operations inherited from $V$. 
+		Let $V$ be a vector space defined over the field of scalars $\mathbb{F}$ and $U \subset V$, $U$ 
+		is subspace if it is a vector space over $\mathbb{F}$ for the operations inherited from $V$. 
 		To be a subspace $U$ must
@@
 	\begin{thm}
-		A nonempty subset $U$ of a vector space $V$ is a subspace if and only if $U$ is closed with respect to addition and scalar multiplication.
+		A non-empty subset $U$ of a vector space $V$ is a subspace if and only if $U$ is closed with respect to addition and scalar multiplication.
 	\end{thm}
@@
 	\begin{thm}
-		Suppose $X$ is a set of vectors from a vector space $V$. The subspace spanned by $X$ is the smallest subspace $W$ of $V$ containing $X$. We write that $W = \left< X \right>$.
+		Suppose $X$ is a set of vectors from a vector space $V$. The subspace spanned by $X$ is the smallest 
+		subspace $W$ of $V$ containing $X$. We write that $W = \left< X \right>$.
 	\end{thm}
@@
 	\begin{defn}
-		A linear combination of the vectors from $X$ is an expression of the form $w = \sum_{i \in I} a_i u_i$ where $a_i \in \mathbb{F}$ (scalars) and $u_i \in X$ (vectors).
-		If the set $X$ is infinite, them only finitely many $a_i$ are allowed to be nonzero. $w$ will of course be a vector, referred to as a linear combination of the $u_i$'s.
+		A linear combination of the vectors from $X$ is an expression of the form $w = \sum_{i \in I} a_i u_i$ 
+		where $a_i \in \mathbb{F}$ (scalars) and $u_i \in X$ (vectors).	If the set $X$ is infinite, them only 
+		finitely many $a_i$ are allowed to be non-zero. $w$ will of course be a vector, referred to as a linear 
+		combination of the $u_i$'s.
 	\end{defn}
@@
 	\begin{thm}
-		Suppose $V$ is a vector space over $\mathbb{F}$, and $X = \{ u_i | i \in I \}$ is a set of vectors in $V$. Then $W=\left< X \right>$ consists of all vectors $w$ that are linear
-		combinations of the vectors from $X$.
+		Suppose $V$ is a vector space over $\mathbb{F}$, and $X = \{ u_i | i \in I \}$ is a set of vectors 
+		in $V$. Then $W=\left< X \right>$ consists of all vectors $w$ that are linear combinations of the vectors from $X$.
 	\end{thm}
@@
 	\subsection{Is a vector in the subspace?}
-	From above theorem, to check a given vector is contained in the subspace $W=\left< X \right>$, just check $w$ is a linear combination of the vectors $u_i$ from $X$.
+	From above theorem, to check a given vector is contained in the subspace $W=\left< X \right>$, just check $w$ is a 
+	linear combination of the vectors $u_i$ from $X$.
 
 \section{Bases and dimension}
+	\subsection{Linear independence}
 	\begin{defn}
+		If a non-zero linear combinatino of the vector from $X$ adds up to the zero vector then we say that $X$ is a linearly dependent set of vectors.
 	\end{defn}
+	Similarly,
+	\begin{defn}
+		The set $X$ is linearly independent if no nonzero linear combination of the vectors from $X$ sums to the zero vector.
+	\end{defn}
+
+	\subsection{Bases}
+	\begin{defn}
+		A basis of $V$ is a linearly indepedent spanning set for $V$.
+	\end{defn}
+
+	\subsection{Minimal spanning sets}
+	\begin{thm}
+		Suppose $X$ and $Y$ are subsets of a vector space $V$ and that $Y \subseteq X$. Then $<X> \subseteq <Y>$.
+	\end{thm}
+
+	\begin{thm}
+		A set of vectors $X$ is a minimal spanning set for $U = <X>$ if and only if $X$ is linearly independent.
+	\end{thm}
+
+	\subsection{Bases always exist}
+	\begin{thm}
+		If $V$ is a vector space then $V$ contains at least one basis.
+	\end{thm}
+
+	\subsection{Size of a linearly independent set}
+	\begin{thm}
+		If a vector space $V$ has a finite basis of size $n$ then every linearly independent set of vectors in $V$ has a size at most $n$.
+	\end{thm}
+
+	\subsection{Dimension}
+	\begin{crl}
+		If $V$ has a finite spanning set then every basis in $V$ has the same finite size.
+	\end{crl}
+
+	\begin{defn}
+		The dimension of a vector space $V$ is the size of an arbitrary basis in it.
+	\end{defn}
+	Note, we say $V$ is finite/infinitely dimensional depending on whether the size of the basis is finite/infinite.
+
+	\begin{thm}
+		A given set of $n$ vectors $v_1, v_2, ..., v_n$ from $\mathbb{F}^n$ is a basis if and only if the square matrix ($n \times n$) with rows $v_i$ has a non-zero determinant.
+	\end{thm}
+
+	\subsection{Coordinates}
 
